## Introduction

This document includes the code for all exercises that were introduced during the First R Training on October 4th 2016. The dataset 'compensation.csv' can be downloaded from: https://github.com/Larnsce/datasets.git

## The first 3 Steps

Step 1 - Clear R's Brain
```{r, echo = TRUE,  message = FALSE, warning = FALSE}

# Always include this line of code at the top your scirpt
rm(list = ls()) 

```

Step 2 - Load needed libraries into R
```{r, echo = TRUE,  message = FALSE, warning = FALSE}

# Load needed libraries into R
library(dplyr) 

```

Step 3 - Read the compensation dataset into R using read.csv()
```{r, echo = FALSE,  message = FALSE, warning = FALSE}

# Read data into R
cmps <- read.csv("/Users/Your Name/Sub-Folder1/Sub-Folder2/0 - datasets/compensation.csv")

```

## Using dplyr for data manipulation

#### Exercise - Slide 66

```{r, echo = TRUE, eval = FALSE,  message = FALSE, warning = FALSE}

# 1. Use a dplyr function to get the 8th row
slice(cmps, 8)

# 2. Use a dplyr function to get the Fruit column
select(cmps, Fruit)

# 3. Use dplyr functions to get the 8th row of the Fruit column
slice(select(cmps, Fruit), 8)

# alternative
select(slice(cmps, 8), Fruit)

# another alternative
FruitOnly <- select(cmps, Fruit)
slice(FruitOnly, 8)

# or "piping"
cmps %>%
  select(Fruit) %>%
  slice(8)

```

#### Logical Operators - Slide 72

```{r, echo = TRUE, eval = FALSE,  message = FALSE, warning = FALSE}

x <- seq(from = 1, to = 9, by = 1)    # makes a sequence from 1 to 9 by steps of 1
y <- rep(5,9)                         # repeats the number 5 nine times
x == y             # checks in which places x and y are the the same and responses with "true" and "false"
x < y                                 # is x smaller than y?
x > y                                 # is x bigger than y?
x >= y                                # is x bigger or equal to y?
x != y                                # is x not equal to y?

```

#### Exercise - Slide 74

```{r, echo = TRUE, eval = FALSE,  message = FALSE, warning = FALSE}

# 1. subset the dataset to only have Fruit values bigger than 80
#     - Hint: use filer()
filter(cmps, Fruit > 80)

# 2. Modify your commands, so that you see only the Root column returned
select(filter(cmps, Fruit > 80), Root)

# alternative: piping
cmps %>%
  filter(Fruit > 80) %>%
  select(Root)

```

#### Exercise Slide 83

```{r, echo = TRUE, eval = TRUE,  message = FALSE, warning = FALSE}
summarise(cmps, 
	meanFruit = mean(Fruit),
	sdFruit = sd(Fruit))

```




